High resolution, annual maps of the characteristics of smallholder-dominated croplands at national scales

### 2021-03-02 inst/manuscript/manuscript.Rmd
Edited Results and SI

#opts_knit$set(root.dir=normalizePath('../'))

@@

##### Mapping assignments
Labellers registered in the system log in to the mapping platform (built with Flask) and navigate to the OpenLayers-based field mapping interface (Figure \ref{fig:labeller}), where they are presented with a white target box representing the randomly selected grid cell, a set of digitizing tools, and different image backdrops, including true and false color renderings of the the growing season and off-growing season PlanetScope composites, and several virtual globe basemaps. Following a set of pre-defined digitizing rules, the labeller uses the polygon drawing tool to digitize the boundaries of all crop fields intersecting the target grid cell that are visible within the PlanetScope overlays. To aid with interpretation, the labeller can toggle between the PlanetScope renderings and the basemaps to help form a judgement about what constitutes a field. The labeller assigns each polygon a field class category (e.g. annual cropland, tree crop, fallow), and upon completing all fields submits the assignment to the database. In cases where the target grid cell does not contain any fields, the labeller simply submits the assignment to mark it complete. The labeller is then directed to the next available assignment from a different labelling task.

@@

To collect the initial randomized samples for model training, we grouped the AOIs into three blocks: a northern block comprising the 6 northernmost AOIs, a central to southeastern blocks comprising the 3 middle (AOIs 7-9) and 2 southeastern AOIs (12 and 15), and a southwestern block made up of the forest zone AOIs (10, 11, 13, 14, 16). Within each block, we randomly selected and labelled 500 grid cells, which provided relatively large initial training samples for these agro-ecologically similar regions, while helping to minimize the overall amount of labelling effort. In addition to these samples, we randomly selected and labelled 100 grid cells within each AOI to provide a validation sample.

After collecting the initial training and validation samples, we trained a starter model for each block and applied it to each of the block's AOI. For each iteration, 100 samples were actively selected within each AOI, and added to the training pool.

@@

### Accuracy assessment
The model performance assessments described above (2.3.2-3) were not fully independent because they used the same validation sites over multiple iterations [@ElmesAccountingtrainingdata2020]. To independently assess the accuracy of our final map products, we followed recommended guidelines [@StehmanKeyissuesrigorous2019] to create a separate map reference sample. To develop this sample, we first defined cropland and non-cropland strata for Ghana by intersecting the centroids of the 0.005 degree grid (Figure \ref{fig:aois}C) with the segmented field boundaries, and then randomly selected samples from the points representing each stratum. The selected sample was loaded into labelling platform, were it was presented as a ~0.1 ha square polygon (representing the minimum mapping unit). An expert supervisor who was not involved in labelling evaluated the PlanetScope composites at each site to determine its class membership. If the 0.1 ha target was not completely contained within a single class, the supervisor drew a square polygon of ~0.1 ha and shifted it the minimum distance necessary to make it contained entirety within the class that the initial target had the greatest overlap. Four classes were used for the map reference sample: cropland; non-cropland; unsure but likely cropland; unsure but likely non-cropland. The latter two classes were used to provide insight into the degree of uncertainty in the map reference sample. For efficiency, two separate supervisors evaluated separate portions of the reference sample, but both assessed a small subset. We calculated their level of agreement on this subset to provide an additional assessment of uncertainty in the map reference sample [@StehmanKeyissuesrigorous2019]. The SI contains further details on the creation and analysis of the map reference sample.

@@

## Assessing the characteristics of Ghanaian cropland
Using the final mapped results, we calculated the estimated area of arable cropland in Ghana, as well as the average size of fields and number of fields in the different AOIs. We used the map reference sample to calculate adjusted area estimates and confidence intervals for each map class, and used the biases between labellers' polygons and segmented boundaries at validation labels to calculate adjusted estimates of mean field sizes and the total number of fields.

# Results
```{r, echo=FALSE, message=FALSE, warning=FALSE}
tiles <- sf::read_sf(system.file("extdata/ghana_tiles.geojson", 
                                 package = "activemapper"))
km2 <- round(as.numeric(units::set_units(sum(sf::st_area(tiles)), "km2")))
km2 <- format(km2, big.mark = ",", scientific = FALSE)
```
We developed maps for all of Ghana (`r km2` km$^2$). Ghana has several agro-ecologically distinct regions, ranging from the primarily grain and vegetable crop producing regions in the northern savannas to tree crop-dominated system in the forested southwest, where cocoa and oil palm are among the dominant crops. For these latter regions, we did not attempt to classify tree crops, and instead mapped clearings that potentially contain field crops or newly felled or replanted tree crops. We made this decision because the resolution of PlanetScope imagery is insufficient to distinguish tree crops from the surrounding forest, and clear field boundaries are often not visible within cocoa crops.

## Image catalog and quality
```{r, echo=FALSE}
data("image_quality")
tqual_mu <- image_quality$tile %>%
  group_by(season, tile) %>% 
  summarize(score = mean(Score)) %>% 
  ungroup()
tqual_sum <- tqual_mu %>% 
  group_by(season) %>% 
  summarize(mu = mean(score), sd = sd(score))
tqmu <- tqual_sum %>% pull(mu) %>% round(2)
names(tqmu) <- tqual_sum %>% pull(season)

tqual_ct <- tqual_mu %>% group_by(season) %>% 
  mutate(gt85 = ifelse(score > 0.85, 1, 0)) %>% 
  summarize(n = n(), n85 = sum(gt85)) %>% 
  mutate(prop = n85 / n)
tqct <- tqual_ct %>% pull(prop) * 100
names(tqct) <- tqual_ct %>% pull(season)
```

We generated a complete image catalog for the `r nrow(tiles)` tiles intersecting Ghana, spanning `r km2` km$^2$. This entailed processing all PlanetScope imagery intersecting the total extent of these tiles for the periods May-September, 2018 (the growing season) and December, 2018 to February, 2019 (the subsequent dry season). The longer period was necessary for the growing season because of the high frequency of cloud cover, which severely limits the number of clear scenes over any given area. For cloudier regions (AOIs 10, 11, 13, 14, 16) we started the dry season window in November. The compositing processes was applied to render the resulting seasonally contrasting image pairs for each tile. We also composited imagery for AOI 8 (Figure \ref{fig:aois}A) for the two seasons of the 2019 agricultural year. All images were converted into Raster Foundry slippy maps for presentation in *labeller*.

An assessment with two observers (see SI for observer details) found that average quality per growing season composite tile was `r tqmu["gs"]`, with `r tqct["gs"]` percent having scores $\geq$ 0.85, while the average quality of dry season composites was `r tqmu["os"]` (`r tqct["os"]` percent $\geq$ 0.85). Composite quality in both seasons was highest in the northern half of the country and lowest in the southwest (Figure \ref{fig:imqual}A), where the substantially greater cloud cover resulted in a much lower density of available PlanetScope imagery for each time period.

@@

```{r imqual, echo = FALSE, out.width="70%", fig.cap="The location and quality scores of 100 randomly selected tiles for the growing (A) and off-growing season (B), and the corresponding distributions of the quality scores for each season, respectively (C and D) [This a hypothetical placeholder figure].", fig.align='center'}
knitr::include_graphics('figures/figure4.png')
```

## Active learning
### Training data collection
```{r, echo=FALSE, message=FALSE}
initial <- train_val_sites$initial %>% as_tibble %>% group_by(Cluster) %>%
  count()
active <- train_val_sites$active %>% as_tibble %>% filter(usage == "train") %>% 
  group_by(aoi) %>% count()
typicaln <- raster::modal(initial$n) + raster::modal(active$n)
aoi10n <- (initial %>% filter(Cluster %in% 3) %>% pull(n)) + 
  (active %>% filter(aoi == "labeller10") %>% pull(n))
aoi14n <- (initial %>% filter(Cluster %in% 3) %>% pull(n)) + 
  (active %>% filter(aoi == "labeller14") %>% pull(n))
aoi15n <- (initial %>% filter(Cluster %in% 2) %>% pull(n)) + 
  (active %>% filter(aoi == "labeller15") %>% pull(n))
aoi3init <- initial %>% filter(!Cluster %in% 1:3) %>% pull(n)
aoi3n <- aoi3init + (active %>% filter(aoi == "labeller3") %>% pull(n))

initial_total <- train_val_sites$initial %>% as_tibble() %>% distinct(name) %>% 
  count()
active_total <- sum(active$n)
validate_total <- train_val_sites$active %>% as_tibble() %>% 
  filter(usage == "validate") %>% distinct(name) %>% count()
```

After initial training, the active learning process was run for 3 iterations for 12 of the 16 AOIs, resulting in `r typicaln` labels per AOI (Figure S2A). AOIs 10 and 14 stopped after one and two iterations, respectively, as they started with high initial validation accuracies (>83%) and showed little subsequent improvement, thus were trained with `r aoi10n` -  `r aoi14n` samples. AOI 15 was run for 4 iterations and had a training sample size of `r aoi15n`, while AOI 3 underwent a second active learning cycle because the model produced during the first cycle was inaccurate (see SI). In this second run, `r aoi3init`  initial training sites randomly selected within the AOI were used (Figure S2A), followed by 2 subsequent active learning subsequent iterations, resulting in a training sample of `r aoi3n`. The total number of unique training and validation sites across the country were `r initial_total + active_total` and `r validate_total`, respectively.

@@

# training rate
train_rate <- label_summary$nlabels %>% 
  tidyr::pivot_wider(names_from = kml_type, values_from = n) %>% 
  rename_all(funs(c("ID", "tr", "tref"))) %>% 
  mutate(trate = tref / tr) %>% filter(tr > 10)
```

The distribution of training and validation sample collection effort was divided across `r nrow(label_summary$score_stats)` labellers, with a core team of `r label_summary$nlabels %>% filter(kml_type == "Training" & n > 1000) %>% nrow()` who mapped more than 1,000 sites each. As each training site was mapped by 4 separate labellers, `r tsites$tot[1]` sets of vector labels were made. Each labeller digitized an average of `r tsites$mu[1]` (see Figure S3A for more details on labelling effort).

This accuracy of the labellers was assessed `r tsites$tot[2]` times against `r length(unique(label_summary$scores$name))` unique training reference sites (Figure S3A), with each labeller assessed an average of `r tsites$mu[2]` times at a rate of 1 training reference site for every `r round(1 / (tsites$mu[2] / tsites$mu[1]), 2)` training site mapped. The average of each labeller's average accuracy score was `r round(mean(label_summary$score_stats$Mean), 2)` (range `r round(min(label_summary$score_stats$Mean), 2)` to `r round(max(label_summary$score_stats$Mean), 2)`; see Figure S3B for detailed score distributions).

@@

After each site was mapped by four labellers, consensus labels were generated. The Bayesian Risk (see SI) of each training and validation label was calculated as an additional measure of label quality. The average risk was `r murisk$Train` for training labels and `r murisk$Validate` for validation labels. Risk was highest in the northern cluster (AOIs 1-6; Figure S5), falling between `r cl1$Risk[1]` for training and `r cl1$Risk[2]` for validation labels (Figure S6), and lowest in the southwestern cluster (AOIs 10, 11, 13, 14, 16; training risk = `r cl3$Risk[1]`; validation risk = `r cl3$Risk[2]`). Label risk in the central/eastern cluster (AOIs 7-9, 12, 15) was slightly lower (training = `r cl2$Risk[1]`; validation = `r cl2$Risk[2]`) than in the north. Greater labeller experience also appeared to reduce risk, as the mean risk for the second mapping of Cluster 2's 500 initial training sites was `r cl2brisk`, compared to `r cl2arisk` for the first mapping.

**[To further evaluate label quality, we assessed the degree of similarity between worker polygons with the field boundaries collected by our ground validation team. Overall, we found that on average a worker's field boundary interpreted from a PlanetScope image had only an *X%* agreement (as determined by area of intersection over union) with the corresponding field-mapped boundary, and the average field size mapped by a worker was on average *X% smaller/larger* than those mapped on the ground.]**

### Validation performance
```{r, echo=FALSE}
data("iteration_metrics")
delta_func <- function(x, rnd = 4) round((x - lag(x)) / lag(x), rnd)

@@

The gain in model performance was assessed for each AOI against several accuracy metrics. The average accuracy, AUC, and F1 at iteration 0 were `r imu0$Accuracy$score`, `r imu0$AUC$score`, and `r imu0$F1$score`, respectively, increasing to `r imu3$Accuracy$score`, `r imu3$AUC$score`, and `r imu3$F1$score` by iteration 3 (Figure \ref{fig:alperformance}). These differences represent respective gains of `r imu3$Accuracy$diff`, `r imu3$AUC$diff`, and `r imu3$F1$diff` percent for the three metrics. The largest gains in each metric were registered on iteration 1, averaging `r idmu$I1$Accuracy$score`, `r idmu$I1$AUC$score`, and `r idmu$I1$F1$score` percent for accuracy, AUC, and F1. The lowest gains were received on Iteration 3 for accuracy (`r idmu$I3$Accuracy$score`%) and F1 (`r idmu$I3$F1$score`%), while AUC gains were least on iteration 2 (`r idmu$I2$AUC$score`%). The range in scores varied substantially across AOIs and metrics. Accuracy ranged between `r imin[imin$Metric == "Accuracy", ]$Score` (AOI `r imin[imin$Metric == "Accuracy", ]$AOI`) and `r imax[imax$Metric == "Accuracy", ]$Score` (AOI `r imax[imax$Metric == "Accuracy", ]$AOI`), while AUC varied from `r imin[imin$Metric == "AUC", ]$Score` (AOI `r imin[imin$Metric == "AUC", ]$AOI`) and `r imax[imax$Metric == "AUC", ]$Score` (AOI `r imax[imax$Metric == "AUC", ]$AOI`), and F1 from `r imin[imin$Metric == "F1", ]$Score` (AOI `r imin[imin$Metric == "F1", ]$AOI`) and `r imax[imax$Metric == "F1", ]$Score` (AOI `r imax[imax$Metric == "F1", ]$AOI`).

```{r alperformance, echo = FALSE, out.width="100%", fig.cap="Scores for overall Accuracy, area under the curve of the Receiver Operating Characteristic, and the F1 score resulting for the Random Forest model after each iteration of the active learning loop for each AOI (gray lines), as well as the mean score per iteration across all AOIs (black lines).", fig.align='center', message=FALSE}
knitr::include_graphics('figures/figure5.png')
```

```{r, echo=FALSE}
# random versus active data
data("iteration_metrics")

@@

The training tests at AOIs 1, 8, and 15 showed that active learning produced higher performance than randomized model training after three iterations, with accuracy, AUC, and F1 being respectively `r u(ixl$III["Accuracy"])`, `r u(ixl$III["AUC"])`, and `r u(ixl$III["F1"])` percent higher after three iterations (Figure S5). However, there was more variability in earlier iterations, with average score differences of `r u(ixl$I["Accuracy"])` (accuracy), `r u(ixl$I["AUC"])` (AUC), and `r u(ixl$I["F1"])` percent (F1) after iteration 1, and `r u(ixl$II["Accuracy"])` (accuracy), `r u(ixl$II["AUC"])` (AUC), and `r u(ixl$II["F1"])` (F1) percent after iteration 2. The lower accuracy under active leaving was caused by results at AOI 15, where accuracy gain was `r abs(u(i1acc["15"]))` percent lower under active training (although accuracy slightly between iterations; see Figure \ref{fig:alperformance}). In comparison, active learning accuracies after iteration 1 were `r u(i1acc["1"])` and `r u(i1acc["8"])` percent higher than random training for AOIs 1 and 8, respectively.

@@

The potential impact of training data errors on map quality was assessed in four AOIs (1, 2, 8, and 15). The results of these tests showed that the average accuracy, AUC, and F1 scores for models trained with the consensus labels were respectively `r metdiffs$Accuracy$cons`, `r metdiffs$AUC$cons`, and `r metdiffs$F1$cons` (Figure \ref{fig:trainingimpact}). Models trained with the most accurate worker's labels had scores that were just `r metdiffs$AUC$chdelt` - `r metdiffs$Accuracy$chdelt` percent higher than models trained with the most accurate worker's lablers (accuracy = `r metdiffs$Accuracy$high`'; AUC = `r metdiffs$AUC$high`; F1 = `r metdiffs$F1$high`), but were `r metdiffs$AUC$cldelt` - `r metdiffs$Accuracy$cldelt` higher than models trained with the least accurate worker's labels (accuracy = `r metdiffs$Accuracy$low`; AUC = `r metdiffs$AUC$low`; F1 = `r metdiffs$F1$low`).

```{r trainingimpact, echo = FALSE, out.width="100%", fig.cap="Scores for overall Accuracy, area under the curve of the Receiver Operating Characteristic, and the F1 score resulting from Random Forest models trained with consensus labels, and labels made by the most and least accurate worker to label each site. Comparisons were made for AOIs 1, 2, 8, and 15, denoted by grey symbols, while the mean scores across these AOIs are shown for each metric.", fig.align='center', message=FALSE}
knitr::include_graphics('figures/figure6.png')
```

```{r, echo=FALSE}
data("label_risk_metrics")

@@

A second measure of the impact of training data error can be obtained by calculating the between the average training label risk per AOI versus various model performance metrics (Table S3). Accuracy had a strong negative association (Spearman's Rank Correlation =  `r risk_corr[, "Accuracy"]` and AUC had a moderate negative association (r = `r risk_corr[, "AUC"]`), while F1 had a weaker positive association (r = `r risk_corr[, "F1"]`). This sign of the latter is counter-intuitive, but relates to the fact that precision was higher in areas of increasing label risk (`r risk_corr[, "Precision"]`), as did false positive error rates (`r risk_corr[, "FPR"]`). In contrast, recall had little correlation with label risk (`r risk_corr[, "Recall"]`), thus the higher precision drove F1 higher, even in areas with more false positives.  **[work on this]**

@@

To assess the overall accuracy of the final maps, we used our map reference sample of `r sum(ref_ct)` sites (`r sum(crp_ct$n)` cropland; `r sum(noncrp_ct$n)` non-cropland) to evaluate the accuracy of 1) pixel-wise classifications produced by the RandomForest model, and 2) the segmented field boundary maps. We first evaluated the uncertainty in the map reference classes using two approaches, the first being an assessment of the overall agreement between map reference labels (at `r sum(usertab)` sites), and the second was the level of confidence in assigned labels (see SI for details). The first measure showed that the two individual labellers agreed `r agreement`% of sites, while the second showed that `r round(sum(crp_ct$n[2], noncrp_ct$n[2]) / sum(ref_ct) * 100, 1)` of sites labelled were of lower confidence.

>>>>

fmt <- function(x) format(x, big.mark = ",")
#opts_knit$set(root.dir=normalizePath('../'))

@@

##### Mapping assignments
Labellers registered in the system log in to the mapping platform (built with Flask) and navigate to the OpenLayers-based field mapping interface (Figure \ref{fig:labeller}), where they are presented with a white target box representing the randomly selected grid cell, a set of digitizing tools, and different image backdrops, including true and false color renderings of the the growing season and off-growing season PlanetScope composites, and several virtual globe basemaps. Following a set of pre-defined digitizing rules (see SI), the labeller uses the polygon drawing tool to digitize the boundaries of all crop fields intersecting the target grid cell that are visible within the PlanetScope overlays. To aid with interpretation, the labeller can toggle between the PlanetScope renderings and the basemaps to help form a judgement about what constitutes a field. The labeller assigns each polygon a field class category (e.g. annual cropland, tree crop, fallow), and upon completing all fields submits the assignment to the database. In cases where the target grid cell does not contain any fields, the labeller simply submits the assignment to mark it complete. The labeller is then directed to the next available assignment from a different labelling task.

@@

To collect the initial randomized samples for model training, we grouped the AOIs into three clusters: a northern cluster comprising the 6 northernmost AOIs (Cluster 1), a central to southeastern cluster (Cluster 2) consisting of the 3 middle (AOIs 7-9) and 2 southeastern AOIs (12 and 15), and a southwestern cluster (Cluster 3) made up of the forest zone AOIs (10, 11, 13, 14, 16). Within each cluster, we randomly selected and labelled 500 grid cells, which provided relatively large initial training samples for these agro-ecologically similar regions, while helping to minimize the overall amount of labelling effort. In addition to these samples, we randomly selected and labelled 100 grid cells within each AOI to provide a validation sample.

After collecting the initial training and validation samples, we trained a starter model for each cluster and applied it to each of the block's AOI. For each iteration, 100 samples were actively selected within each AOI, and added to the training pool.

@@

### Accuracy assessment
The model performance assessments described above (2.3.2-3) were not fully independent because they used the same validation sites over multiple iterations [@ElmesAccountingtrainingdata2020]. To independently assess the accuracy of our final map products, we followed recommended guidelines [@StehmanKeyissuesrigorous2019] to create a separate map reference sample. We used a stratified design, randomly assigning square polygons of ~0.1 ha extent into cropland and non-cropland strata, developed the map of segmented field boundaries. Four classes were used for the map reference sample: cropland; non-cropland; unsure but likely cropland; unsure but likely non-cropland. The latter two classes were used to provide insight into the degree of uncertainty in the map reference sample. For efficiency, two separate supervisors evaluated separate portions of the reference sample, but both jointly assessed a small subset of the sample. We calculated their level of agreement on this subset to provide an additional assessment of uncertainty in the map reference sample [@StehmanKeyissuesrigorous2019]. The SI contains further details on the design and collection of the map reference sample.

@@

## Assessing the characteristics of Ghanaian cropland
Using the final mapped results, we calculated the estimated area of arable cropland in Ghana, as well as the average size and total number of fields in the different AOIs. We used the map reference sample to calculate adjusted area estimates and confidence intervals for each map class, and used the differences between labellers' polygons and segmented boundaries at validation sites to calculate bias-adjusted estimates of mean field sizes and the total number of fields.

# Results
```{r, echo=FALSE, message=FALSE, warning=FALSE}
tiles <- sf::read_sf(system.file("extdata/ghana_tiles.geojson", 
                                 package = "activemapper"))
km2 <- round(as.numeric(units::set_units(sum(sf::st_area(tiles)), "km2")))
km2 <- fmt(km2)
```
We developed maps of Ghana's arable croplands, covering an area of `r km2` km$^2$ that included portions of neighboring countries overlapped by images tiles. Ghana has several agro-ecologically distinct regions, ranging from the primarily grain and vegetable crop producing regions in the northern savannas to tree crop-dominated system in the forested southwest, where cocoa and oil palm are among the dominant crops. For these latter regions, we did not attempt to classify tree crops, and instead mapped clearings that potentially contain field crops or newly felled or recently replanted tree crops. We made this decision because PlanetScope's resolution is not high enough for labellers to distinguish many tree crops from surrounding forest, and the boundaries of many tree crops (e.g. cocoa) are often not visible.

## Image catalog and quality
```{r, echo=FALSE, message=FALSE, warning=FALSE}
data("image_quality")
tqual_mu <- image_quality$tile %>%
  group_by(season, tile) %>% 
  summarize(score = mean(Score)) %>% 
  ungroup()
tqual_sum <- tqual_mu %>% 
  group_by(season) %>% 
  summarize(mu = mean(score), sd = sd(score))
tqmu <- tqual_sum %>% pull(mu) %>% round(2)
names(tqmu) <- tqual_sum %>% pull(season)

tqual_ct <- tqual_mu %>% group_by(season) %>% 
  mutate(gt85 = ifelse(score > 0.85, 1, 0)) %>% 
  summarize(n = n(), n85 = sum(gt85)) %>% 
  mutate(prop = n85 / n)
tqct <- tqual_ct %>% pull(prop) * 100
names(tqct) <- tqual_ct %>% pull(season)

nt <- fmt(nrow(tiles))
```

To develop the map, we first generated an image catalog for the `r nt` tiles covering Ghana. This entailed processing all PlanetScope imagery intersecting these tiles between May-September, 2018 (the growing season) and December, 2018 to February, 2019 (the subsequent dry season). The longer period was necessary for the growing season because of the frequent cloud cover, which severely limits the number of clear scenes for any tile. For the cloudiest regions (AOIs 10, 11, 13, 14, 16) we started the dry season window in November.

An assessment with two observers (see SI for observer details) found that average quality per growing season composite tile was `r tqmu["Growing"]`, with `r tqct["Growing"]` percent having scores $\geq$ 0.85, while the average quality of dry season composites was `r tqmu["Dry"]` (`r tqct["Dry"]` percent $\geq$ 0.85). Composite quality in both seasons was highest in the northern half of the country and lowest in the southwest (Figure \ref{fig:imqual}A), where the substantially greater cloud cover resulted in a much lower density of available PlanetScope imagery for each time period.

@@

```{r imqual, echo = FALSE, out.width="70%", fig.cap="The location and quality scores of 100 randomly selected tiles for the growing (A) and off-growing season (B), and the corresponding distributions of the quality scores for each season, respectively (C and D).", fig.align='center'}
knitr::include_graphics('figures/figure4.png')
```

## Active learning
### Training data collection
```{r, echo=FALSE, message=FALSE, warning=FALSE}
initial <- train_val_sites$initial %>% as_tibble %>% group_by(Cluster) %>%
  count()
active <- train_val_sites$active %>% as_tibble %>% filter(usage == "train") %>% 
  group_by(aoi) %>% count()
typicaln <- raster::modal(initial$n) + raster::modal(active$n)
aoi10n <- (initial %>% filter(Cluster %in% 3) %>% pull(n)) + 
  (active %>% filter(aoi == "labeller10") %>% pull(n))
aoi14n <- (initial %>% filter(Cluster %in% 3) %>% pull(n)) + 
  (active %>% filter(aoi == "labeller14") %>% pull(n))
aoi15n <- (initial %>% filter(Cluster %in% 2) %>% pull(n)) + 
  (active %>% filter(aoi == "labeller15") %>% pull(n))
aoi3init <- initial %>% filter(!Cluster %in% 1:3) %>% pull(n)
aoi3n <- aoi3init + (active %>% filter(aoi == "labeller3") %>% pull(n))

initial_total <- train_val_sites$initial %>% as_tibble() %>% distinct(name) %>% 
  count()
active_total <- sum(active$n)
validate_total <- train_val_sites$active %>% as_tibble() %>% 
  filter(usage == "validate") %>% distinct(name) %>% count()

ttrain <- fmt(initial_total + active_total)
tval <- fmt(validate_total$n)
```

After training of models with the initial randomly selected label sets, the active learning process was run for 3 iterations for 12 of the 16 AOIs, resulting in `r typicaln` labels per AOI (see Figure S3A for sample distributions). AOIs 10 and 14 stopped after one and two iterations, respectively, as they started with high initial validation accuracies (>83%) and showed little subsequent improvement. The models for these two AOIs were thus trained with `r aoi10n` -  `r aoi14n` samples. AOI 15 was run for 4 iterations (`r aoi15n` samples), while AOI 3 underwent a second active learning cycle because the model produced during the first cycle was inaccurate (see SI). In this second run, `r aoi3init` initial training sites randomly selected within the AOI were used (Figure S3A), followed by 2 subsequent active learning iterations, resulting in a training sample of `r aoi3n`. The total number of unique training and validation sites across the country were `r ttrain` and `r tval`, respectively.

@@

# training rate
train_rate <- label_summary$nlabels %>% 
  tidyr::pivot_wider(names_from = kml_type, values_from = n) %>% 
  rename_all(funs(c("ID", "tr", "tref"))) %>% 
  mutate(trate = tref / tr) %>% filter(tr > 10)

tsitesc <- tsites %>% mutate(across(where(is.numeric), fmt))
# nvecs <- tsites$tot[1])
# muvecs <- fmt(tsites$mu[1])

The distribution of training and validation sample collection effort was divided across `r nrow(label_summary$score_stats)` labellers, with a core group of `r label_summary$nlabels %>% filter(kml_type == "Training" & n > 1000) %>% nrow()` who mapped more than 1,000 sites each. As each training site was mapped by 4 separate labellers, `r tsitesc$tot[1]` sets of vector labels were made. Each labeller digitized an average of `r tsitesc$mu[1]` (see Figure S4A for more details on labelling effort). Labeller accuracy was scored `r tsitesc$tot[2]` times against `r length(unique(label_summary$scores$name))` unique training reference sites (Figure S3A), with each labeller assessed an average of `r tsites$mu[2]` times at a rate of 1 training reference site for every `r round(1 / (tsites$mu[2] / tsites$mu[1]), 2)` training site mapped. The mean of each labeller's average accuracy score was `r round(mean(label_summary$score_stats$Mean), 2)` (range `r round(min(label_summary$score_stats$Mean), 2)` to `r round(max(label_summary$score_stats$Mean), 2)`; see Figure S4B for detailed score distributions).

@@

After each site was mapped by four labellers, consensus labels were generated. The Bayesian Risk (see SI) of each training and validation label was calculated as an additional measure of label quality. The average risk was `r murisk$Train` for training labels and `r murisk$Validate` for validation labels. Risk was highest in the northen AOIs (AOIs 1-6; Figures S5-6), falling between `r cl1$Risk[1]` for training and `r cl1$Risk[2]` for validation labels (Figures S5-6), and lowest in the southwestern AOIs (AOIs 10, 11, 13, 14, 16; training risk = `r cl3$Risk[1]`; validation risk = `r cl3$Risk[2]`). Label risk in the central-southeastern AOIs (AOIs 7-9, 12, 15) was slightly lower (training = `r cl2$Risk[1]`; validation = `r cl2$Risk[2]`) than in the north. Labeller experience also appeared to reduce risk, which we observed during a relabelling of the 500 initial random site in this cluster (see SI); the mean risk of the updated labels was `r cl2brisk`, compared to `r cl2arisk` for original labels.

### Validation performance
```{r, echo=FALSE, warning=FALSE, message=FALSE}
data("iteration_metrics")
delta_func <- function(x, rnd = 4) round((x - lag(x)) / lag(x), rnd)

@@

Model performance was calculated for each iteration within each AOI. The average accuracy, AUC, and F1 at iteration 0 were `r imu0$Accuracy$score`, `r imu0$AUC$score`, and `r imu0$F1$score`, respectively, increasing to `r imu3$Accuracy$score`, `r imu3$AUC$score`, and `r imu3$F1$score` by iteration 3 (Figure \ref{fig:alperformance}). These differences represent respective gains of `r imu3$Accuracy$diff`, `r imu3$AUC$diff`, and `r imu3$F1$diff` percent for the three metrics. The largest gains for each metric occurred on iteration 1, averaging `r idmu$I1$Accuracy$score`, `r idmu$I1$AUC$score`, and `r idmu$I1$F1$score` percent for accuracy, AUC, and F1, while the lowest gains were realized on iteration 3, with accuracy, F1, and AUC respectively increasing by just `r idmu$I3$Accuracy$score`%, `r idmu$I3$F1$score`%, and `r idmu$I2$AUC$score`%. The scores achieved on the final iteration varied substantially across AOIs and metrics. Accuracy ranged between `r imin[imin$Metric == "Accuracy", ]$Score` (AOI `r imin[imin$Metric == "Accuracy", ]$AOI`) and `r imax[imax$Metric == "Accuracy", ]$Score` (AOI `r imax[imax$Metric == "Accuracy", ]$AOI`), while AUC varied from `r imin[imin$Metric == "AUC", ]$Score` (AOI `r imin[imin$Metric == "AUC", ]$AOI`) and `r imax[imax$Metric == "AUC", ]$Score` (AOI `r imax[imax$Metric == "AUC", ]$AOI`), and F1 from `r imin[imin$Metric == "F1", ]$Score` (AOI `r imin[imin$Metric == "F1", ]$AOI`) and `r imax[imax$Metric == "F1", ]$Score` (AOI `r imax[imax$Metric == "F1", ]$AOI`).

```{r alperformance, echo = FALSE, out.width="100%", fig.cap="Scores for overall Accuracy, area under the curve of the Receiver Operating Characteristic, and the F1 scores for the Random Forests model results after each iteration of the active learning loop for each AOI (gray lines), as well as the mean score per iteration across all AOIs (black lines).", fig.align='center', message=FALSE}
knitr::include_graphics('figures/figure5.png')
```

```{r, echo=FALSE, message=FALSE, warning=FALSE}
# random versus active data
data("iteration_metrics")

@@

The comparison of active versus randomized training sample collection (in AOIs 1, 8, and 15) showed that the former approach outperformed the latter. After three iterations, the accuracy, AUC, and F1 scores resulting from active learning were respectively `r u(ixl$III["Accuracy"])`, `r u(ixl$III["AUC"])`, and `r u(ixl$III["F1"])` percent higher than the scores from a randomly trained model (Figure S7). However, there was more variability in earlier iterations, with average score differences of `r u(ixl$I["Accuracy"])` (accuracy), `r u(ixl$I["AUC"])` (AUC), and `r u(ixl$I["F1"])` percent (F1) after iteration 1, and `r u(ixl$II["Accuracy"])` (accuracy), `r u(ixl$II["AUC"])` (AUC), and `r u(ixl$II["F1"])` (F1) percent after iteration 2. The negative results for accuracy was caused by results at AOI 15, where active learning accuracy was `r abs(u(i1acc["15"]))` percent lower than random training after iteration 1 (see Figure \ref{fig:alperformance}). In comparison, iteration 1 active learning accuracies 1 were `r u(i1acc["1"])` and `r u(i1acc["8"])` percent higher than random training for AOIs 1 and 8, respectively. Accuracy under active learning for AOI 15 exceeded randomized training after 4 iterations.

@@

The potential impact of label errors on map quality was assessed in four AOIs (1, 2, 8, and 15). The results of these tests showed that the average accuracy, AUC, and F1 scores for models trained with the consensus labels were respectively `r metdiffs$Accuracy$cons`, `r metdiffs$AUC$cons`, and `r metdiffs$F1$cons` (Figure \ref{fig:trainingimpact}). Performance metrics from consensus-trained models were just `r metdiffs$AUC$chdelt` - `r metdiffs$Accuracy$chdelt` percent higher than those models trained with the most accurate individuals' labels (accuracy = `r metdiffs$Accuracy$high`'; AUC = `r metdiffs$AUC$high`; F1 = `r metdiffs$F1$high`), but were `r metdiffs$AUC$cldelt` - `r metdiffs$Accuracy$cldelt` higher than models trained with the least accurate individual labels (accuracy = `r metdiffs$Accuracy$low`; AUC = `r metdiffs$AUC$low`; F1 = `r metdiffs$F1$low`).

```{r trainingimpact, echo = FALSE, out.width="100%", fig.cap="Scores for overall accuracy, area under the curve of the Receiver Operating Characteristic, and the F1 score resulting from models trained with consensus labels, and labels made by the most and least accurate labellers to map each site. Comparisons were made for AOIs 1, 2, 8, and 15, denoted by grey symbols, while the mean scores across these AOIs are shown for each metric.", fig.align='center', message=FALSE}
knitr::include_graphics('figures/figure6.png')
```

```{r, echo=FALSE, message=FALSE, warning=FALSE}
data("label_risk_metrics")

@@

A second measure of the impact of label error is found within the correlations between the mean label risk per AOI and the model performance metrics (Table S3). Accuracy and AUC had strong (Spearman's Rank Correlation =  `r risk_corr[, "Accuracy"]` to moderate (r = `r risk_corr[, "AUC"]`) negative correlations with label risk, while F1 had a weaker but moderate positive association (r = `r risk_corr[, "F1"]`). The positive sign of the latter relationship is counter-intuitive, but is explained by risk's association with precision, one of two inputs to F1, which was moderately positive (`r risk_corr[, "Precision"]`), whereas risk had a negligible correlation with recall (`r risk_corr[, "Recall"]`), F1's other component. The correlation between risk and the false positive rate (`r risk_corr[, "FPR"]`), another important performance metric, shows that labelling uncertainty may increase model commission error.

@@

We used a map reference sample of `r sum(ref_ct)` sites (`r sum(crp_ct$n)` cropland; `r sum(noncrp_ct$n)` non-cropland) to evaluate the accuracy of the per-pixel classifications (resulting from thresholding the Random Forests probability), as well as the segmented field boundary maps. We first evaluated the uncertainty in the map reference classes by assessing 1) the overall agreement between map reference labels collected by two separate supervisors at `r sum(usertab)` sites, and 2) the confidence of the labels assigned by the supervisors (see SI for details). The first measure showed that the two individual supervisors' labels agreed at `r agreement`% of common sites, while the second showed that `r round(sum(crp_ct$n[2], noncrp_ct$n[2]) / sum(ref_ct) * 100, 1)` of sites were labelled with the two classes that indicated a level of uncertainty.
